 ---
title: "Strings: In-class Exercises (Part 2)"
format:
  html: default
editor_options: 
  chunk_output_type: console
editor: 
  markdown: 
    wrap: 72
---

You can download this .qmd file from
[here](https://github.com/joeroith/264_spring_2026/blob/main/03_strings_part2.qmd).
Just hit the Download Raw File button.

This uses parts of R4DS Ch 14: Strings and Ch 15: Regular Expressions
(both the first and second editions).

## Manipulating strings

str functions to know for manipulating strings:

-   str_length(): Count how many characters are there in the string 
-   str_sub(): Pull out character base on the pocision
-   str_c(): to combine all the character string  
-   str_to_lower(): everything to  lowercase 
-   str_to_upper(): Everything upper case 
-   str_to_title():make the first letter capital 
-   str_replace() *not in 02_strings examples*

```{r}
#| message: false
#| warning: false

library(tidyverse)

#spotify <- read_csv("Data/spotify.csv") 
spotify <- read_csv("spotify.csv")


spot_smaller <- spotify |>
  select(
    title, 
    artist, 
    album_release_date, 
    album_name, 
    subgenre, 
    playlist_name
  )

spot_smaller <- spot_smaller[c(5, 32, 49, 52, 83, 175, 219, 231, 246, 265), ]
spot_smaller
```

## Warm-up

0.  Describe what EACH of the str\_ functions below does. Then, create a
    new variable "month" which is the two digit month from
    album_release_date

```{r}
spot_new <- spot_smaller |>
  select(title, album_release_date) |>
  mutate(title_length = str_length(title), #count how many character in the title 
         year = str_sub(album_release_date, 1, 4),
         month = parse_number(str_sub(album_release_date, 6,7)), #Pull out the 2-degit month as a string  
         title_lower = str_to_lower(title), #make all the letter become lower case 
         album_release_date2 = str_replace_all(album_release_date, "-", "/")) #replace the character (-) in the variable album_release_date with (/)
spot_new
         
max_length <- max(spot_new$title_length)

str_c("The longest title is", max_length, "characters long.", sep = " ") #combine all the characters inside the quatation mark where the max_length is computing the longest length of the song title 
str_glue("The longest title is {max_length} characters long.") #does the same thing as str_c but just use {} to reduce using ""
```

## Important functions for identifying strings which match

str_view() : most useful for testing str_subset() : useful for printing
matches to the console str_detect() : useful when working within a
tibble

1.  Identify the input type and output type for each of these examples:

```{r}
#str_view(spot_smaller$subgenre, "pop") #input : character vector, output: an htmlwidget object
#typeof(str_view(spot_smaller$subgenre, "pop")) #input:htmlwidget object, output: character string can be a list  
#class(str_view(spot_smaller$subgenre, "pop")) #input: htmlwidget object, output: character vector of class labels

#str_view(spot_smaller$subgenre, "pop", match = NA) #input: character vector, output: htmlwidget object(with different highlighting behavior because of match = NA)
#str_view(spot_smaller$subgenre, "pop", html = TRUE) #character vector, output: htmlwidget object 

str_subset(spot_smaller$subgenre, "pop") #input: character vector, output: character vector 

str_detect(spot_smaller$subgenre, "pop") #input: string vector, output is the logical vector, only True or False 
```

2.  Use str_detect to print the rows of the spot_smaller tibble
    containing songs that have "pop" in the subgenre. (i.e. make a new
    tibble with fewer rows)
```{r}
spot_smaller |>
  filter(str_detect(spot_smaller$subgenre, "pop"))
  
```
    

3.  Find the mean song title length for songs with "pop" in the subgenre
    and songs without "pop" in the subgenre.
```{r}
spot_smaller|>
  mutate(sub_pop = str_detect(spot_smaller$subgenre, "pop"),
         title_length = str_length(title))|>
  group_by(sub_pop)|>
  summarize(mean_title_length = mean(title_length), .group= "drop")  #if it's False - no title pop in it 
    #if true- the title pop is there 
```
    
    
  


Producing a table like this would be great:

```{r}
#| eval: false

# A tibble: 2 × 2
  sub_pop mean_title_length
  <lgl>               <dbl>
1 FALSE                18.6
2 TRUE                 13.6
```

Producing a table like this would be
SUPER great (hint: `ifelse()`):
```{r}
spot_smaller|>
  mutate(sub_pop2 = ifelse(str_detect(spot_smaller$subgenre, "pop"),
                           "Genre with pop",
      "Genre without pop"),
         title_length = str_length(spot_smaller$title))|>
  group_by(sub_pop2)|>
  summarize(mean_title_length = mean(title_length), .groups = "drop")
 
```


```{r}
#| eval: false

# A tibble: 2 × 2
  sub_pop           mean_title_length
  <chr>                         <dbl>
1 Genre with pop                 13.6
2 Genre without pop              18.6
```

4.  In the bigspotify dataset, find the proportion of songs which
    contain "love" in the title (track_name) by playlist_genre.

```{r}
#| message: false
#| warning: false

bigspotify <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

bigspotify

str_view(str_to_lower(bigspotify$track_name), "love")

bigspotify|>
  filter(!is.na(track_name))|>
  mutate(title_love = str_to_lower(track_name),
         love = str_detect(title_love, "love"))|>
  group_by(playlist_genre)|>
  summarize(pop_love = mean(love),
            count =n())
```
```

```

## Matching patterns with regular expressions

|          |                                          |
|----------|------------------------------------------|
| `^abc`   | string starts with exactly "abc"         |
| `abc$`   | string ends with exactly "abc"           |
| `.`      | any character (helpful as a placeholder) |
| `[^abc]` | anything EXCEPT a or b or c              |

```{r}
# Guess the output!

#str_view(spot_smaller$artist, "^K") #start with K

#str_view(spot_smaller$album_release_date, "01$") #end with 01

#str_view(spot_smaller$title, "^.. ") #any first 2 characters 

#str_view(spot_smaller$artist, "[^A-Za-z ]") #any artist who don't have regular letter in it 
```



5.  Given the corpus of common words in stringr::words, create regular
    expressions that find all words that:
```{r, eval=FALSE}
#str_view(stringr :: words, "y") # Start with “y”.
#str_view(stringr :: words, "x$") # End with “x”
#str_view(stringr :: words, "^...$") #Are exactly three letters long.
#str_view(stringr :: words, "^.......")# Have seven letters or more.
#str_view(stringr :: words, "^[A,E,I,O,U,a,e,i,o,u]")# Start with a vowel.
#str_view(stringr :: words, "^[^e]ed$")# End with ed, but not with eed.
#str_view(stringr :: words, "^q[^u]")# Words where q is not followed by u. (are there any in `words`?)
```
  


```{r}
# Try using str_view() or str_subset()

# For example, to find words with "tion" at any point, I could use:
#str_view(words, "tion") #shows where the strings is
str_subset(words, "tion")
```

## More useful regular expressions:

|       |                              |
|-------|------------------------------|
| `\\d` | any number                   |
| `\\s` | any space, tab, etc.         |
| `\\b` | any boundary: space, ., etc. |

```{r}
#str_view(spot_smaller$album_name, "\\d") #any degit 
#str_view(spot_smaller$album_name, "\\s") #any space
#str_view(spot_smaller$album_name, "\\b") #any boundary: space
```

Here are the regular expression special characters that require an
escape character (a preceding `\`): `^` `$` `.` `?` `*` `|` `+` `(` `)`
`[` `{`

For any characters with special properties, use `\` to "escape" its
special meaning ... but `\` is itself a special character ... so we need
two `\\`! (e.g. `\\$`, `\\.`, etc.)

```{r}
#str_view(spot_smaller$title, "$")
#str_view(spot_smaller$title, "\\$") #if you want to find the $ you gotta use \\$
```

6.  In bigspotify, how many track_names include a `$`? Be sure you print
    the track_names you find and make sure the dollar sign is not just
    in a featured artist!
```{r}
bigspotify|>
  filter(str_detect(track_name, "\\$"))|>
  filter(!str_detect(track_name, "(feat | with).*\\$"))|> #if you do [feat] it will mean you don't want 
  select(track_name, track_artist)|>
  print(n = Inf)
```
    
    
7.  In bigspotify, how many track_names include a dollar amount (a `$`
    followed by a number).
```{r}
bigspotify|>
  filter(str_detect(track_name, "\\$\\d"))|>
  select(track_name, track_artist)
```

or 

```{r}
bigspotify|>
  filter(str_detect(track_name, "\\$[0-9]"))|>
  select(track_name, track_artist)
```
## Repetition

|         |                       |
|---------|-----------------------|
| `?`     | 0 or 1 times          |
| `+`     | 1 or more             |
| `*`     | 0 or more             |
| `{n}`   | exactly n times       |
| `{n,}`  | n or more times       |
| `{,m}`  | at most m times       |
| `{n,m}` | between n and m times |



```{r}
#str_view(spot_smaller$album_name, "[A-Z]{2,}")
#str_view(spot_smaller$album_release_date, "\\d{4}-\\d{2}")
         
```


**Use at least 1 repetition symbol when solving 8-10 below**

8.  Modify the first regular expression above to also pick up "A.A" (in
    addition to "BEYONC" and "II"). That is, pick up strings where there
    might be a period between capital letters.
    
```{r}
#str_view(spot_smaller$album_name, "[A-Z]{2,}| [A-Z]\\.[A-Z]+") #the + can capture the captial letter after that 
  
  
```

or 
```{r}
#str_view(spot_smaller$album_name, "[A-Z.]{2,}") # the dot inside the [] , capital letter followed by dot 
```

9.  Create some strings that satisfy these regular expressions and
    explain.

-   `"^.*$"` ( Start with any character, end with any character)


-   `"\{.+\}"` (looking for the string that have {}, that have something inside of it)
```{r}
#str_view("This is $200", "^.*$ ") # start with anything, ends with anything - haven' escaped the $
#str_view("Why is this I {that}",`"\{.+\}"`  ) # anything that includes {}, anything amount of degit or characters that appear 1 or more time 
```

10. Create regular expressions to find all `stringr::words` that:

-   Start with three consonants.
```{r}
#str_view(stringr :: words, "^[^aeiou]{3}")
```

-   Have two or more vowel-consonant pairs in a row.
```{r}
#str_view(stringr :: words, "([aeiou] | [^^aeiou] |[aeiou] | [^^aeiou]){2,} ")
#str_view(stringr :: words, "([aeiou] | [^^aeiou] |[aeiou] | [^^aeiou]){2,} ")
```


## Useful functions for handling patterns

str_extract() : extract a string that matches a pattern str_count() : ( spring detect )
count how many times a pattern occurs within a string

```{r}
str_extract(spot_smaller$album_release_date, "\\d{4}-\\d{2}")


spot_smaller |>
  select(album_release_date) |>
  mutate(year_month = str_extract(album_release_date, "\\d{4}-\\d{2}"))

spot_smaller |>
  select(artist) |>
  mutate(n_vowels = str_count(artist, "[aeiou]"))
```

11. In the spot_smaller dataset, how many words are in each title?
    (hint: `\\b`)
```{r}
spot_smaller |>
  select(title)|>
  mutate(title2 = str_count(title,"\\b\\w+\\b" )) #w+ means let the words repeat 
```
    

12. In the spot_smaller dataset, extract the first word from every
    title. Show how you would print out these words as a vector and how
    you would create a new column on the spot_smaller tibble. That is,
    produce this:
```{r}

str_extract(spot_smaller$title, "^[^ ]+")


str_sub(str_extract(spot_smaller$title, "^[^ ]+"))

```
    
```{r}
spot_smaller|>
  select(title)|>
  mutate(one_word_title = str_extract(spot_smaller$title, "^[^ ]+"))
 
```

```{r}
# [1] "Hear"      "Run"       "Formation" "7/11"      "My"        "It's"     
# [7] "Poetic"    "A.D.H.D"   "Ya"        "Runnin"   
```

Then this:

```{r}
# A tibble: 10 × 2
#   title                                             first_word
#   <chr>                                             <chr>     
# 1 Hear Me Now                                       Hear      
# 2 Run the World (Girls)                             Run       
# 3 Formation                                         Formation 
# 4 7/11                                              7/11      
# 5 My Oh My (feat. DaBaby)                           My        
# 6 It's Automatic                                    It's      
# 7 Poetic Justice                                    Poetic    
# 8 A.D.H.D                                           A.D.H.D   
# 9 Ya Estuvo                                         Ya        
#10 Runnin (with A$AP Rocky, A$AP Ferg & Nicki Minaj) Runnin    
```

13. Which decades are popular for playlist_names? Using the bigspotify
    dataset, try doing each of these steps one at a time!

-   filter the bigspotify dataset to only include playlists that include
    something like "80's" or "00's" in their title.
-   create a new column that extracts the decade
-   use count to find how many playlists include each decade
-   what if you include both "80's" and "80s"?
-   how can you count "80's" and "80s" together in your final tibble?

```{r}
bigspotify|>
  select(playlist_name, track_name)|>
  filter(str_detect(playlist_name, "\\d{2}'s")| 
        str_detect(playlist_name, "\\d{2}s"))|>
  mutate(decade = str_extract(playlist_name,"\\d{2}'s|\\d{2}s"))|>
  count(decade, sort = TRUE)
  
  
```


  

## Grouping and backreferences

```{r}
# find all fruits with repeated pair of letters.  
fruit = stringr::fruit
fruit
#str_view(fruit, "(..)\\1", match = TRUE) #2 same characters follow by each other 

# why does the code below add "pepper" and even "nectarine"?
#str_view(fruit, "(..)(.*)\\1", match = TRUE)# any single chracter, follow by single character, 
```

(...) position 
Tips with backreference:

-   You must use () around the the thing you want to reference.
-   To backreference multiple times, use `\\1` again.
-   The number refers to which spot you are referencing... e.g. `\\2`
    references the second set of ()

```{r}
x1 <- c("abxyba", "abccba", "xyaayx", "abxyab", "abcabc")
#str_view(x1, "(.)(.)(..)\\2\\1")
#str_view(x1, "(.)(.)(..)\\1\\2")
#str_view(x1, "(.)(.)(.)\\1\\2\\3")
```

14. Describe to your groupmates what these expressions will match, and
    provide a word or expression as an example:
    
```{r}
fruit = stringr::fruit
#str_view(stringr::words, "(.)\\1\\1", match = TRUE)

```
    

-   `"(.)\\1\\1"` : This will capture the tripple repeated letters
-   `"(.)(.)(.).*\\3\\2\\1"`: First 3 letters, mirrors the end, abccab  

Which words in `stringr::words` match each expression?


15. Construct a regular expression to match words in `stringr::words`
    that contain a repeated pair of letters (e.g. “church” contains “ch”
    repeated twice) but *not* match repeated pairs of numbers (e.g.
    507-786-3861).
```{r}
#str_view(stringr::words, "([A-Za-z]{2}).*\\1", match = TRUE) 
```
    

16. Reformat the album_release_date variable in spot_smaller so that it
    is MM-DD-YYYY instead of YYYY-MM-DD. (Hint: str_replace().)

```{r}
spot_smaller|>
  select(album_release_date)|>
  mutate(album_release_date =
           str_replace(album_release_date,
                       "^(\\d{4})-(\\d{2})-(\\d{2})$",
                       "\\2-\\3-\\1"))
  
```

17. BEFORE RUNNING IT, explain to your partner(s) what the following R
    chunk will do:
     **Capture the first second and thrid string that come after the space. The first string/ character would stay the same but the 2nd and 3rd change possition with each other** 

```{r}
sentences |> 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") |> 
  head(5)
```
