---
title: "HW1"
format: html
---

```{r}
library(readr)
vaccinations_2021 <- read_csv("vaccinations_2021.csv")


```

```{r}
#| include: false
library(tidyverse)
vaccinations_2021 <- read_csv("vaccinations_2021.csv")

```

```{r}
list.files()

```



1.  Approach 1: create a Data folder in the same location where this
    .qmd file resides, and then store vaccinations_2021.csv in that Data
    folder
2.  Approach 2: give R the complete path to the location of
    vaccinations_2021.csv, starting with Home (\~)
3.  Approach 3: link to our course webpage, and then know we have a Data
    folder containing all our csvs
4.  Approach 4: navigate to the data in GitHub, hit the Raw button, and
    copy that link

A recent Stat 272 project examined determinants of covid vaccination
rates at the county level. Our data set contains 3053 rows (1 for each
county in the US) and 14 columns; here is a quick description of the
variables we'll be using:

-   `state` = state the county is located in
-   `county` = name of the county
-   `region` = region the state is located in
-   `metro_status` = Is the county considered "Metro" or "Non-metro"?
-   `rural_urban_code` = from 1 (most urban) to 9 (most rural)
-   `perc_complete_vac` = percent of county completely vaccinated as of
    11/9/21
-   `tot_pop` = total population in the county
-   `votes_Trump` = number of votes for Trump in the county in 2020
-   `votes_Biden` = number of votes for Biden in the county in 2020
-   `perc_Biden` = percent of votes for Biden in the county in 2020
-   `ed_somecol_perc` = percent with some education beyond high school
    (but not a Bachelor's degree)
-   `ed_bachormore_perc` = percent with a Bachelor's degree or more
-   `unemployment_rate_2020` = county unemployment rate in 2020
-   `median_HHincome_2019` = county's median household income in 2019

```{r}
library(dplyr)
library(ggplot2)
library(forcats)

```

1.  Consider only Minnesota and its surrounding states (Iowa, Wisconsin,
    North Dakota, and South Dakota). We want to examine the relationship
    between the percentage who voted for Biden and the percentage of
    complete vaccinations by state. Generate two plots to examine this
    relationship:

<!-- -->

a)  A scatterplot with points and smooters colored by state. Make sure
    the legend is ordered in a meaningful way, and include good labels
    on your axes and your legend. Also leave off the error bars from
    your smooters.

what you need for this - call in the data - filter by state -ggplot
-geom_point -geom_smooth -mutate fct_reorder

```{r}
vaccinations_2021|>
  filter(state %in% c( "Minnesota","Iowa", "Wisconsin", "North Dakota", "South Dakota"))|>
  ggplot(aes(x = perc_Biden,
             y = perc_complete_vac, 
             color= fct_reorder2(state, perc_complete_vac, perc_Biden))) +
  geom_point()+
  geom_smooth(se = FALSE) +
  labs(title = " The relationship between percentage of county voting for Biden and the covid 19 vaccination rate in the country",
  x = "% voting for Biden",
  y= "% fully vaccinated",
  color = "state")
```

b)  One plot per state containing a scatterplot and a smoother.

```{r}
vaccinations_2021 |>
   filter(state %in% c( "Minnesota","Iowa", "Wisconsin", "North Dakota", "South Dakota"))|>
  ggplot(aes(x = perc_Biden,
             y = perc_complete_vac, 
             color= fct_reorder2(state, perc_complete_vac, perc_Biden))) +
  geom_point()+
  geom_smooth(se = FALSE) +
  labs(title = " The relationship between percentage of county voting for Biden and the covid 19 vaccination rate in the country",
  x = "% voting for Biden",
  y= "% fully vaccinated",
  color = "state") +
  facet_wrap(~state)
  
  
```

Describe which plot you prefer and why. What can you learn from your
preferred plot?

***Second one because it's easier for me to see the trend clearly in
every state***

2.  We wish to compare the proportions of counties in each region with
    median household income above the national median (\$69,560).

<!-- -->

a)  Fill in the blanks below to produce a segmented bar plot with
    regions ordered from highest proportion above the median to lowest.

```{r}
library(tidyverse)

income <- vaccinations_2021 |>
  mutate(HHincome_vs_national = ifelse(median_HHincome_2019 >= 69560,
                                       "Above median",
                                       "Below Median"))

region_order <- income |>
  group_by(region) |>
  summarise(prop_above = mean(median_HHincome_2019 >= 69560, na.rm = TRUE)) |>
  arrange(desc(prop_above)) |>
  pull(region)

income |>
  mutate(region_sort = fct_relevel(region, region_order)) |>
  ggplot(aes(x = region_sort, fill = HHincome_vs_national)) +
  geom_bar(position = "fill") +
  labs(x = "Region", y = "Proportion of counties", fill = "Income vs national median")

```

b)  Create a table of proportions by region to illustrate that your bar
    plot in (a) is in the correct order (you should find two regions
    that are *really* close when you just try to eyeball differences).

```{r}
income |>
  group_by(region) |>
  summarise(
    n = n(),
    prop_above = mean(median_HHincome_2019 >= 69560, na.rm = TRUE)
  ) |>
  arrange(desc(prop_above))

```

c)  Explain why we can replace `fct_relevel(region, FILL IN CODE)` with

`mutate(region_sort = fct_reorder(region, median_HHincome_2019 < 69560, .fun = mean))`

but not

`mutate(region_sort = fct_reorder(region, median_HHincome_2019 < 69560))`

```{r}
#| eval: false

vaccinations_2021 |>
  mutate(HHincome_vs_national = ifelse(median_HHincome_2019 < 69560, "Below median", "Above median")) |>
  mutate(region_sort = fct_relevel(region, "Northeast", "West", "South")) |>
  mutate(region_sort = fct_reorder(region,median_HHincome_2019 < 69560 ))|>
  ggplot(mapping = aes(x = region_sort, fill = HHincome_vs_national)) +
    geom_bar(position = "fill")
```

3.  We want to examine the distribution of total county populations and
    then see how it's related to vaccination rates.

<!-- -->

a)  *Carefully and thoroughly* explain why the two histograms below
    provide different plots.

```{r}
vaccinations_2021 |>
  mutate(tot_pop_millions = tot_pop / 1000000) |>
  ggplot(mapping = aes(x = tot_pop_millions)) +
    geom_histogram(bins = 40) +
    labs(x = "Total population in millions")


vaccinations_2021 |>
  mutate(tot_pop_millions = tot_pop %/% 1000000) |>
  ggplot(mapping = aes(x = tot_pop_millions)) +
    geom_histogram(bins = 40) +
    labs(x = "Total population in millions")
```

*The first one uses regular division, producing a continuous measure of
population in millions, while the second uses integer division, which
truncates values and forces counties into whole-number population
categories.*

b)  Find the top 5 counties in terms of total population.

```{r}
vaccinations_2021 |>
  arrange(desc(tot_pop)) |>
  slice_head(n = 5) |>
  select(state, county, tot_pop)

```

c)  Plot a histogram of logged population and describe this
    distribution.

```{r}
vaccinations_2021 |>
 mutate(log_poptot = log(tot_pop))|>
   ggplot( aes(x = log_poptot)) +
    geom_histogram(bins = 40) 
   

  

```

d)  Plot the relationship between log population and percent vaccinated
    using separate colors for Metro and Non-metro counties (be sure
    there's no 3rd color used for NAs). Reduce the size and transparency
    of each point to make the plot more readable. Describe what you can
    learn from this plot.

```{r}
vaccinations_2021 |>
  filter(!is.na(metro_status)) |>
  mutate(log_pop = log10(tot_pop)) |>
  ggplot(aes(x = log_pop,
             y = perc_complete_vac,
             color = metro_status)) +
  geom_point(size = 1, alpha = 0.3) +
  geom_smooth(se = FALSE) +
  labs(
    x = "log10(Total county population)",
    y = "% fully vaccinated",
    color = "Metro status"
  )

```

4.  Produce 3 different plots for illustrating the relationship between
    the rural_urban_code and percent vaccinated. Hint: you can sometimes
    turn numeric variables into categorical variables for plotting
    purposes (e.g. `as.factor()`, `ifelse()`).

```{r}
ggplot(vaccinations_2021,
       aes(x = rural_urban_code, y = perc_complete_vac)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Rural–Urban Code (1 = most urban, 9 = most rural)",
    y = "% fully vaccinated",
    title = "Rural-Urban Code vs Percent of county completely vaccinated as of 11/9/21"
    
  )

```

```{r}
vaccinations_2021 |>
  mutate(ruc = as.factor(rural_urban_code)) |>
  ggplot(aes(x = ruc, y = perc_complete_vac)) +
  geom_boxplot() +
  labs(
    x = "Rural–Urban Code",
    y = "% fully vaccinated", 
    title = " Rural-Urban Code vs Percent of county completely vaccinated as of 11/9/21 "
  )

```

**This is a boxplot showing the distribution of the percent of county
population fully vaccinated as of 11/9/21 by rural–urban code. X axis
represent rural–urban code ranging from 1 = most urban to 9 = most
rural. The Y axis represents percentage of county completely vaccinated.
Each box represents counties within a rural–urban code category, and
points indicate outliers. he general pattern shows that median
vaccination rates are highest for the most urban counties (codes 1 and
2). Overall, the plot suggests that counties that are more rural tend to
have lower and more variable vaccination rates than more urban
counties.**

```{r}
vaccinations_2021|>

  mutate(urban_group = ifelse(rural_urban_code <= 3,
                              "Urban (1–3)",
                              "Rural (4–9)")) |>
  ggplot(aes(x = urban_group, y = perc_complete_vac)) +
  geom_violin() +
  labs(
    x = "",
    y = "% fully vaccinated",
    title = "Rural-Urban Code vs Percent of county completely vaccinated as of 11/9/21 "
  )

  
```

State your favorite plot, why you like it better than the other two, and
what you can learn from your favorite plot. Create an alt text
description of your favorite plot, using the Four Ingredient Model. See
[this
link](https://drive.google.com/file/d/1BuEbL6vETBcFULmtlrwwZBUYhdEg0Zqb/view?usp=sharing)
for reminders and references about alt text.

5.  BEFORE running the code below, sketch the plot that will be produced
    by R. AFTER running the code, describe what conclusion(s) can we
    draw from this plot?

    **It is a scatter plot. it is ordered by states from the lowest
    interquatile range of proportion voting for Biden to the highest.
    IQR_Biden is on the x axis and States name on the Y-axis**

```{r}
#| eval: FALSE

vaccinations_2021 |>
  filter(!is.na(perc_Biden)) |>
  mutate(big_states = fct_lump(state, n = 10)) |>
  group_by(big_states) |>
  summarize(IQR_Biden = IQR(perc_Biden)) |>
  mutate(big_states = fct_reorder(big_states, IQR_Biden)) |>
  ggplot() + 
    geom_point(aes(x = IQR_Biden, y = big_states))
```

\*\* 6. In this question we will focus only on the 12 states in the
Midwest (i.e. where region == "Midwest").

a)  Create a tibble with the following information for each state. Order
    states from least to greatest state population.

-   number of different `rural_urban_code`s represented among the
    state's counties (there are 9 possible)
-   total state population
-   proportion of Metro counties
-   median unemployment rate

```{r}
vaccinations_2021|>
  filter(region == "Midwest")|>
  group_by(state)|>
  summarize( n_ruc = n_distinct(rural_urban_code),
             state_pop = sum(tot_pop),
            prop_metro = mean(metro_status == "Metro"),
            median_unemp = median(unemployment_rate_2020))|>
  arrange(state_pop)
  

```

b)  Use your tibble in (a) to produce a plot of the relationship between
    proportion of Metro counties and median unemployment rate. Points
    should be colored by the number of different `rural_urban_code`s in
    a state, but a single linear trend should be fit to all points. What
    can you conclude from the plot?

    **States with moderate Metro proportions tend to have higher
    unemployment, but the pattern is not strictly linear and does not
    appear very strong**

```{r}
vaccinations_2021|>
  filter(region == "Midwest")|>
  group_by(state)|>
  summarize( n_ruc = n_distinct(rural_urban_code),
             state_pop = sum(tot_pop),
            prop_metro = mean(metro_status == "Metro"),
            median_unemp = median(unemployment_rate_2020))|>
 ggplot(aes(x = prop_metro,
            y = median_unemp, 
            color = n_ruc)) +
  geom_point(alpha = 0.5, size = 2.5)+
  geom_smooth(se = FALSE)+
  labs(
    x = "Proportion of the county considered Metro ",
    y = " Median unempleyment",
    color = "# of Rural-Urban Codes",
    title = " Relationship between Proportion of county Metro and Median unemployment in 2020"
  )
  
```

7.  Generate an appropriate plot to compare vaccination rates between
    two subregions of the US: New England (which contains the states
    Maine, Vermont, New Hampshire, Massachusetts, Connecticut, Rhode
    Island) and the Upper Midwest (which, according to the USGS,
    contains the states Minnesota, Wisconsin, Michigan, Illinois,
    Indiana, and Iowa). What can you conclude from your plot?

In this next section, we consider a few variables that could have been
included in our data set, but were NOT. Thus, you won't be able to write
and test code, but you nevertheless should be able to use your knowledge
of the tidyverse to answer these questions.

Here are the hypothetical variables:

-   HR_party = party of that county's US Representative (Republican,
    Democrat, Independent, Green, or Libertarian)
-   people_per_MD = number of residents per doctor (higher values =
    fewer doctors)
-   perc_over_65 = percent of residents over 65 years old
-   perc_white = percent of residents who identify as white

8.  Hypothetical R chunk #1:

```{r}
#| eval: false

# Hypothetical R chunk 1
temp <- vaccine_data |>
  mutate(new_perc_vac = ifelse(perc_complete_vac > 95, NA, perc_complete_vac),
         MD_group = cut_number(people_per_MD, 3)) |>
  group_by(MD_group) |>
  summarise(n = n(),
            mean_perc_vac = mean(new_perc_vac, na.rm = TRUE),
            mean_white = mean(perc_white, na.rm = TRUE))
```

a)  Describe the tibble `temp` created above. What would be the
    dimensions? What do rows and columns represent?

    **- 3 x4**

    **- 3 rows of doctor access groups and 4 columns of group label,
    mean vaccination, mean % white and counts**

b)  What would happen if we replaced
    `new_perc_vac = ifelse(perc_complete_vac > 95, NA, perc_complete_vac)`
    with
    `new_perc_vac = ifelse(perc_complete_vac > 95, perc_complete_vac, NA)`?

**You keep counties with vaccination \>95. All the counties \<= 95 will
become NA**

c)  What would happen if we replaced
    `mean_white = mean(perc_white, na.rm = TRUE)` with
    `mean_white = mean(perc_white)`?

**If any county in a group that has perc_white = NA, the whole group
will become NA**

d)  What would happen if we removed `group_by(MD_group)`?

**The summerize() will do to the whole data set, instead of doing it
within the group**

9.  Hypothetical R chunk #2:

```{r}
#| eval: false

# Hypothetical R chunk 2
ggplot(data = vaccine_data) +
  geom_point(mapping = aes(x = perc_over_65, y = perc_complete_vac, 
                           color = HR_party)) +
  geom_smooth()

temp <- vaccine_data |>
  group_by(HR_party) |>
  summarise(var1 = n()) |>
  arrange(desc(var1)) |>
  slice_head(n = 3)

vaccine_data |>
  ggplot(mapping = aes(x = fct_reorder(HR_party, perc_over_65, .fun = median), 
                       y = perc_over_65)) +
    geom_boxplot()
```

a)  Why would the first plot produce an error? **Because the variable
    perc_over_65 doesn't exist in the main dataset**

b)  Describe the tibble `temp` created above. What would be the
    dimensions? What do rows and columns represent? **Since the at the
    end we got slice_head(n = 3), which mean keep the top 3 parties with
    the most counties. I think the dimension is 3 x 2**

**- 3 rows of top 3 parties** **- 2 columns of HR_party and var1**

c)  What would happen if we replaced
    `fct_reorder(HR_party, perc_over_65, .fun = median)` with
    `HR_party`? **The boxplots will be ordered by its default, can be
    alphabetical or whatever the dataset uses, not by the median of
    parc_over_65**.

<!-- -->

10. Hypothetical R chunk #3:

```{r}
#| eval: false

# Hypothetical R chunk 3
vaccine_data |>
  filter(!is.na(people_per_MD)) |>
  mutate(state_lump = fct_lump(state, n = 4)) |>
  group_by(state_lump, rural_urban_code) |>
  summarise(mean_people_per_MD = mean(people_per_MD)) |>
  ggplot(mapping = aes(x = rural_urban_code, y = mean_people_per_MD, 
      colour = fct_reorder2(state_lump, rural_urban_code, mean_people_per_MD))) +
    geom_line()
```

a)  Describe the tibble piped into the ggplot above. What would be the
    dimensions? What do rows and columns represent?

**The dimmension is going to be 5 x 3**

**5 rows: 4 of them are the name of the top largest states, 1 of of them
called " Others" will dedicate to other state beside the 4 states**

**3 columns: State, rural_urban_code and mean_people_per_MD**

b)  Carefully describe the plot created above.

    **It is a multiple lines plots, x-axis= rural_urban_code which 1 is
    the most urban → 9 most rural. Y-axis = mean_people_per_MD which
    represent average people per doctor. Each line Each line shows how
    doctor access changes across rural–urban codes for that state group,
    therefore, we get different colors line. The `fct_reorder2` wil
    change the legend based on how their lines behave at higher x-values
    (usually based on where the line ends).**

c)  What would happen if we removed `filter(!is.na(people_per_MD))`?

    **Any missing value for variable `people_per_MD` will shows up and
    we will get another line called NA and `geom_line()` will produce
    broken lines / gaps**

d)  What would happen if we replaced
    `fct_reorder2(state_lump, rural_urban_code, mean_people_per_MD)`
    with `state_lump`?

    **The legend order becomes the default factor order (often
    alphabetical / original), instead of being ordered by the line “end”
    behavior.**
