---
title: "another_mental_health_distruction "
format: html
---
loading parkage

```{r}
# Load the packages needed for data cleaning, web requests,
# Census data access, and JSON handling
library(tidyverse)
library(stringr)
library(httr)
library(polite)
library(tidycensus)
library(dplyr)
library(jsonlite)
```

**We would like to analyse with this data set the relationship between people’s political views/affiliation across different regions of the country (Midwest, East Coast, etc), and the type of charitable/political organizations they donate to frequently.** 

In this file we're going to extract data from the Census.gov website based on state and region in order too merge with the cps_data.

First check if the website allow us to scrape
```{r}
# Check whether scraping is allowed on these websites
robotstxt::paths_allowed("https://api.census.gov/data/2023/cps/volunteer/sep")
robotstxt::paths_allowed("https://catalog.data.gov")

# create data dictionary
dir.create("data", showWarnings = FALSE)


```

Second: We will create the region base on tidy census
```{r}

# Use the built-in fips_codes dataset from tidycensus
# Keep only the state information we need and remove duplicates 
us_states_level <- fips_codes |> 
  select(state, state_code, state_name) |> 
  distinct() |>
  mutate(
  
    region = case_when(
      state_code %in% c("09", "23", "25", "33", "44", "50") ~ "Northeast",  # CT, ME, MA, NH, RI, VT
      state_code %in% c("34", "36", "42") ~ "Northeast",  # NJ, NY, PA
      state_code %in% c("17", "18", "26", "39", "55") ~ "Midwest",  # IL, IN, MI, OH, WI
      state_code %in% c("19", "20", "27", "29", "31", "38", "46") ~ "Midwest",  # IA, KS, MN, MO, NE, ND, SD
      state_code %in% c("10", "11", "12", "13", "24", "37", "45", "51", "54") ~ "South",  # DE, DC, FL, GA, MD, NC, SC, VA, WV
      state_code %in% c("01", "05", "21", "22", "28", "47") ~ "South",  # AL, AR, KY, LA, MS, TN
      state_code %in% c("40", "48") ~ "South",  # OK, TX
      state_code %in% c("04", "08", "16", "30", "32", "35", "49") ~ "West",  # AZ, CO, ID, MT, NV, NM, UT
      state_code %in% c("02", "06", "15", "41", "53", "56") ~ "West",  # AK, CA, HI, OR, WA, WY
      TRUE ~ NA_character_
    )
  )
us_states_level
```

Now we will create the donation and vlunteer data from census data 
```{r}
 # To see if our new key is listed 
usethis::edit_r_environ("project")
census_api_key <- Sys.getenv("CENSUS_API_KEY")
print(census_api_key) # see if it works
```

Since all the variables are in special codes and it is confusing, we're trying to create a tibble to see what all the variables mean using this url for variable directory(https://api.census.gov/data/2023/cps/volunteer/sep/variables.json) and it is there on the list when you click (https://api.census.gov/data/2023/cps/volunteer/sep)
```{r}
# Get all variable names
vars_meta1 <- fromJSON("https://api.census.gov/data/2023/cps/volunteer/sep/variables.json")
var_names <- names(vars_meta1$variables) 
var_lookup1 <- tibble(
  variable = var_names,
  label = map_chr(vars_meta1$variables, ~ .x$label %||% NA_character_)
)
var_lookup1
```

Here are the variable we found these variables and what are these variables code mean using this link as a guide ( https://www2.census.gov/programs-surveys/cps/techdocs/cpssep23.pdf)
- State = GESTFIPS
- Age = PRTAGE ( 0-79 age in year, 80 as for 80-84 and 85 for 85+)
- Gender = PESEX ( 1 for male and 2 for female)
- Race = PTDTRACE ( 1 = "White",2 = "Black/African American",3 = "American Indian/Alaska Native",4 = "Asian only",5 = "Hawaiian/Pacific Islander Only",6 = "White-Black",7 = "White-AI",8 = "White-Asian",9 = "White-HP")
- Donate to non-political org = PES18 (-9 No answer, -3 Refusal,-2 Do not know,-1 Not in universe,1 Yes,2 No )
- Donate to political org = PES17 (-9 No answer, -3 Refusal,-2 Do not know,-1 Not in universe,1 Yes,2 No)
-How often volunteer = PES16D ( -9 No answer,-3 Refusal,-2 Do not know,-1 Not in universe,1 Basically every   day,2 A few times a week, 3 A few times a month,4 Once a month,5 Less than once a month,6 Not at all)
-Volunteer status = PRSUPVOL (-1 not in universe,1 Volunteer, 2 Not a volunteer)
-Hour volunteer = PTS16E (-9 No answer,-3 Refused,-2 Don't know,-1 Not in universe)
-Volunteer in children activities = PES16A (-9 No answer,-3 Refused,-2 Don't know,-1 Not in universe)

 get the url from census, then create table

```{r}
usethis::edit_r_environ("project")
census_api_key <- Sys.getenv("CENSUS_API_KEY")

# Get URL that includes all of the variables we need
main_url <- str_c(
  "https://api.census.gov/data/2023/cps/volunteer/sep",
  "?get=GESTFIPS,PRTAGE,PESEX,PTDTRACE,PES18,PES17,PES16D,PRSUPVOL,PTS16E,PES16A",
  "&key=", census_api_key
)

census_data3 <- GET(main_url)

if (status_code(census_data3) != 200) {
  stop("API request failed with status: ", status_code(census_data3))
}

details <- content(census_data3, as = "text", encoding = "UTF-8") |>
  jsonlite::fromJSON()

details[[1]]   
details[[2]]

GESTFIPS <- character()
PRTAGE <- double()
PESEX <- double()
PTDTRACE <- double()
PES18 <- double()
PES17 <- double()
PES16D <- double()
PRSUPVOL <- double()
PTS16E <- double()
PES16A <- double()

for(i in 2:length(details)) {
  GESTFIPS[i-1] <- details[[i]][[1]][1]
  PRTAGE[i-1] <- details[[i]][[2]][1]
  PESEX[i-1] <- details[[i]][[3]][1]
  PTDTRACE[i-1] <- details[[i]][[4]][1]
  PES18[i-1] <- details[[i]][[5]][1]
  PES17[i-1] <- details[[i]][[6]][1]
  PES16D[i-1] <- details[[i]][[7]][1]
  PRSUPVOL[i-1] <- details[[i]][[8]][1]
  PTS16E[i-1] <- details[[i]][[9]][1]
  PES16A[i-1] <- details[[i]][[10]][1]
}

census_data3 <- tibble(
  state_code = GESTFIPS,
  age = as.numeric(PRTAGE),
  gender = as.numeric(PESEX),
  race_code = as.numeric(PTDTRACE),
  donate_nonpolitical_org = as.numeric(PES18),
  donate_political_org = as.numeric(PES17),
  volunteer_freq = as.numeric(PES16D),
  volunteer_status = as.numeric(PRSUPVOL),
  hours_volunteered = as.numeric(PTS16E),
  vol_children_activities = as.numeric(PES16A)
)

census_data3
```

```{r}
usethis::edit_r_environ("project")
census_api_key <- Sys.getenv("CENSUS_API_KEY")
print(census_api_key)
url <- str_c(
  "https://api.census.gov/data/2023/cps/volunteer/sep",
  "?get=GESTFIPS,PRTAGE,PESEX,PTDTRACE,PES18,PES17,PES16D,PRSUPVOL,PTS16E,PES16A",
  "&key=", census_api_key
)

census_data1 <- GET(url)
details <- content(census_data1, "parsed")

details[[1]]   
details[[2]]   

GESTFIPS <- character()
PRTAGE <- double()
PESEX <- double()
PTDTRACE <- double()
PES18 <- double()
PES17 <- double()
PES16D <- double()
PRSUPVOL <- double()
PTS16E <- double()
PES16A <- double()
# Creating for loop using api class activity 
for(i in 2:length(details)) {
  GESTFIPS[i-1] <- details[[i]][[1]][1]
  PRTAGE[i-1] <- details[[i]][[2]][1]
  PESEX[i-1] <- details[[i]][[3]][1]
  PTDTRACE[i-1] <- details[[i]][[4]][1]
  PES18[i-1] <- details[[i]][[5]][1]
  PES17[i-1] <- details[[i]][[6]][1]
   PES16D[i-1] <- details[[i]][[7]][1]
  PRSUPVOL[i-1] <- details[[i]][[8]][1]
  PTS16E[i-1] <- details[[i]][[9]][1]
  PES16A[i-1] <- details[[i]][[10]][1]
}

census_data1 <- tibble(
  state_code = GESTFIPS,
  age = parse_number(PRTAGE),
  gender = parse_number(PESEX),
  race_code = parse_number(PTDTRACE),
  donate_nonpolitical_org = parse_number(PES18),
  donate_political_org = parse_number(PES17),
  volunteer_freq = parse_number(PES16D),
  volunteer_status = parse_number(PRSUPVOL),
  hours_volunteered = parse_number(PTS16E),
  vol_children_activities = parse_number(PES16A)
)

census_data1
```

```{r}
```


```{r}
usethis::edit_r_environ("project")
census_api_key <- Sys.getenv("CENSUS_API_KEY")
print(census_api_key)
url <- str_c(
  "https://api.census.gov/data/2023/cps/volunteer/sep",
  "?get=GESTFIPS,PRTAGE,PESEX,PTDTRACE,PES18,PES17,PES16D,PRSUPVOL,PTS16E,PES16A",
  "&key=", census_api_key
)

census_data1 <- GET(url)
details <- content(census_data1, "parsed")

details[[1]]   
details[[2]]   

GESTFIPS <- character()
PRTAGE <- double()
PESEX <- double()
PTDTRACE <- double()
PES18 <- double()
PES17 <- double()
PES16D <- double()
PRSUPVOL <- double()
PTS16E <- double()
PES16A <- double()
# Creating for loop using api class activity 
for(i in 2:length(details)) {
  GESTFIPS[i-1] <- details[[i]][[1]][1]
  PRTAGE[i-1] <- details[[i]][[2]][1]
  PESEX[i-1] <- details[[i]][[3]][1]
  PTDTRACE[i-1] <- details[[i]][[4]][1]
  PES18[i-1] <- details[[i]][[5]][1]
  PES17[i-1] <- details[[i]][[6]][1]
  PES16D[i-1] <- details[[i]][[7]][1]
  PRSUPVOL[i-1] <- details[[i]][[8]][1]
  PTS16E[i-1] <- details[[i]][[9]][1]
  PES16A[i-1] <- details[[i]][[10]][1]
}

census_data1 <- tibble(
  state_code = GESTFIPS,
  age = parse_number(PRTAGE),
  gender = parse_number(PESEX),
  race_code = parse_number(PTDTRACE),
  donate_nonpolitical_org = parse_number(PES18),
  donate_political_org = parse_number(PES17),
  volunteer_freq = parse_number(PES16D),
  volunteer_status = parse_number(PRSUPVOL),
  hours_volunteered = parse_number(PTS16E),
  vol_children_activities = parse_number(PES16A)
)

census_data1
```

```{r}
census_data1 <- census_data1 |>
  mutate(
    
    gender = recode(gender,
                    `1` = "Male",
                    `2` = "Female",
                    .default = "Unknown"),
    race_code = recode(race_code,
                       `1` = "White",
                       `2` = "Black/African American",
                       `3` = "American Indian/Alaska Native",
                       `4` = "Chinese",
                       `5` = "Japanese",
                       `6` = "Other Asian/Pacific Islander",
                       `7` = "Other race",
                       `8` = "Two major races",
                       `9` = "Three or more races",
                       .default = "Unknown"),
    volunteer_status = recode(volunteer_status, 
                              `-1` = "Not in universe",
                               `1` = "Volunteer",
                               `2` = " Volunteer"),
    hours_volunteered = recode(hours_volunteered, 
                              `-9`= "No answer", 
                               `3`= "Refused",
                               `2`= "Don't know", 
                              `-1`= "Not in universe")
    
  )
```


```{r}

```



