

---
title: |
  `\vspace{-3cm}`{=latex}
  SDS264: Knowledge Quiz 1
subtitle: "Strings and Regular Expressions"
author: "Name:_________________"
format:
  pdf: default
editor_options: 
  chunk_output_type: console
---

Guidelines:

-   You may use one hand-written notesheet, turn in with quiz.

Grading:

-   You will receive maximum credit for writing complete code using correct functions and regular expressions. Ex. *`data |> filter(str_detect(words, "[^e]ed$"))`*

-   You will receive partial credit for identifying the correct function and/or correct regular expressions, even if the code is incomplete. So try to at least write these, even if you're not sure how to put it all together. Ex. *Use `str_detect()` and regular expression `"[^e]ed$"`*

Pledge:

I pledge my honor that on this quiz I have neither given nor received assistance not explicitly approved by the professor and that I an aware of no dishonest work.

-   Write your name here to acknowledge the pledge: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

-   OR

-   place an X here if you intentionally are not signing the pledge: \_\_\_\_\_\_\_

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)
library(rvest)
library(tidytext)

# this will have to reflect the path to your data
park_data <- read_csv("~/Data Science 2/project1/park_data_KQ1.csv")
```

## National Park Data

`park_data` is a 54x3 tibble containing information scraped from national park webpages for a past SDS264 final project. A few notes about the 3 columns:

-   `park_code` is a 4-letter code assigned to each National Park used as a key when merging files

-   `address` is comprised of 4 pieces (described from *left* to *right*): the street address (followed by at least 2 spaces), the city, a 2-letter state abbreviation (followed by a comma), and the zip code (which could be 5 digits or 5 digits, a dash, and 4 more digits)

-   `activities` is a string of activities offered at each park, where activities are separated by commas

\newpage

Here's a small preview of `park_data`:

```{r}
print(park_data[c(2,3,4,29,36),])
```

## Quiz Questions

Please answer the following questions using your knowledge of strings and regular expressions. Use `stringr` functions wherever possible.

1.  Find the subset of all `address` entries that contain the word(s) "Headquarters", "Visitor", or "Center" that may or may not be capitalized.

Hint: `str_subset(park_data$activities, "FILL THIS IN!")`

```{r}
park_data |>
  mutate(address_title = str_to_title(address)) |>
  pull(address_title) |>
  str_subset("Headquarters|Visitor|Center")
```



\vspace{3cm}

2.  Create a new tibble containing only national parks in Florida (FL) and Utah (UT).
```{r}
park_data|>
  filter(str_detect(address, "FL | UT"))
```


\vspace{3cm}

3.  The following code creates a new tibble with the `address` column and then the 4 new columns below, which break apart `address` into its 4 pieces: `street_address`, `city`, `state`, and `zip_code`.


**For 3 of the 4 pieces,** explain what each line is doing, and how it produces the desired outcome. (Hint: str_trim() simply trims off any spaces at the beginning or end of a string)

```{r}
four_pieces <- park_data |>
  mutate(
    street_address = str_trim(str_extract(address, "^.*( ){2,}")), # Piece 1
    city = str_trim(str_extract(address, "( ){2,}.*[A-Z]{2}")),    # Piece 2
    city = str_replace(city, " [A-Z]{2}", ""),                     # Piece 2
    state = str_extract(address, "[A-Z]{2},"),                     # Piece 3
    state = str_replace(state, ",", ""),                           # Piece 3
    zip_code = str_extract(address, "[\\d-]+$")                    # Piece 4
  ) |>
  select(address, street_address, city, state, zip_code)
```

```{r}
print(four_pieces, n = 5)
```

Use your new tibble from (3) to answer Questions (4) and (5).

4.  Print the subset of `street_address` entries where there is a street number with an ordinal suffix (e.g. 1st, 54th, 103rd, etc.).
```{r}
four_pieces |>
  filter(str_detect(street_address, "\\d+(st|nd|rd|th"))|>
   select(address, street_address)
```

```{r}
four_pieces |>
   select(street_address)|>
  filter(str_detect(street_address, "\\d+(st|nd|rd|th)"))
  
```

\vspace{3cm}

5.  Arrange `city` names from longest to shortest.
```{r}
four_pieces|>
  select(address , street_address)|>
  filter(str_detect(street_address, "\\d+(st|nd|rd|th)"))
```


\vspace{3cm}

Use the `activities` column from `park_data` to answer Questions (6) - (7). An example of the first entry in that column is give below:

```{r}
cat(str_wrap(park_data$activities[1], width = 80))
```

6.  Create a new column in `park_data` which records the total number of activities in each park. Hint: since activities are separated by commas, the total number of activities can be thought of as "\# commas + 1".
```{r}
park_data |>
  select(park_code, activities)|>
  mutate(activities_num = str_count(activities, ",") + 1)
```

\vspace{3cm}

7.  Find the subset of entries from the `activities` column that have the same consonant appear twice in a row.

```{r}
park_data |>
  filter(str_detect(activities,"([^aeiouAEIOU])\\1")) |>
  select(activities)
```



\vspace{3cm}

----------------

Bonus Question (you do not need to complete): Find a subset of all the alliterative activities (e.g. Fly Fishing, Wildlife Watching, etc.).

Hint: Be careful of commas, you don't want to select things like "Biking, Boating"

```{r}
park_data |>
  mutate(activities = str_replace_all(activities, ",", "")) |>
  filter(str_detect(activities,("\\b([A-Za-z])[A-Za-z]*\\s+\\1[A-Za-z]*"))
                    )
```

```{r}
park_data |>
  separate_rows(activities, sep = ", ") |>
  filter(str_detect(activities,
      ("\\b([A-Za-z])[A-Za-z]*\\s+\\1[A-Za-z]*")))
```

