---
title: "Knowledge Quiz 2"
subtitle: "Sokmean Poch "
format:
  pdf: default
editor_options: 
  chunk_output_type: console
---


Please answer the following questions, render a pdf, and submit via Moodle by 11 PM on Thursday, Mar 26.

Guidelines: 

- No consulting with anyone else (other than clarification questions from Professor Roith)
- You may use only materials from this class (our class webpage, links on Moodle, our 3 online textbooks, files posted to the RStudio server)
- No online searches or use of large language models like ChatGPT

Pledge: 

I pledge my honor that on this quiz I have neither given nor received assistance not explicitly approved by the professor and that I an aware of no dishonest work.

- type your name here to acknowledge the pledge: _____sokmean poch____________________
- OR
- place an X here if you intentionally are not signing the pledge: _______

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(nycflights13)
```


1. Here's is a crazy list that tells you some stuff about data science.  Give code that will produce **exactly** the following outputs. (There are four parts to this question.)

```{r}
kq2_data <- list(
  DRY = c("Do not", "Repeat", "Yourself"),
  knowledge = c("Knowledge", "is", "Power"),
  nyc = flights,
  integers = 1:25,
  doubles = sqrt(1:25),
  toppings = list(sauce = c("marinara", "bbq", "creamy garlic"),
               cheese = c("mozzarella", "goat", "feta"),
               meat = c("pepperoni", "sausage", "chicken", "none"),
               veggie = c("peppers", "onions", "mushrooms", "pineapple"),
               hold = "no"),
  lostpts_to_perc = cbind(pts_lost = 0:50, 
                          perc = (50 - 0:50) / 50)
)

str(kq2_data)
```

a) 

[1] "Power"     "is"        "Knowledge"

```{r}
kq2_data[[2]][c(3,2,1)]
kq2_data$knowledge[c(3,2,1)]
```

b) 
# A tibble: 6 × 3
# Groups:   dest [1]
  dest  carrier mean_arr_delay
  <chr> <chr>            <dbl>
1 MSP   9E                8.09
2 MSP   DL                4.04
3 MSP   EV               10.5 
4 MSP   MQ                9.56
5 MSP   OO               -2   
6 MSP   UA               -5.5 

```{r}
kq2_data[[3]]|>
  filter(dest == "MSP")|>
  group_by(dest, carrier)|>
  summarize(mean_arr_delay = mean(arr_delay, na.rm = TRUE))


```

c) Note: It must be one string and you must subset from kq2_data for the words "bbq", "goat", "chicken", and "mushroom" 

[1] "Give me a bbq pizza with goat cheese, chicken, and mushrooms."
```{r}
paste(
  "Give me a ",
  kq2_data[[6]][[1]][2],
  " pizza with ",
  kq2_data[[6]][[2]][2],
  " cheese, ",
  kq2_data[[6]][[3]][3],
  ", and ",
  kq2_data[[6]][[4]][3],
  sep = " "
)
```

d) 
[1] 10.0  0.8

```{r}
kq2_data[[7]][11, ]
```


2. Write a function called `test_genre_popularity()` that produces a single ggplot. It should allow a user to input:

  1) a tibble
  2) two song genres
  3) and a plot option to create either a boxplot or faceted histograms
  
The title of your plot should indicate the two genres being compared. The subtitle should provide the *unique* test statistic and p-value for the specific genres being used, these values should change in the subtitle when you use different genres in the function input (Hint: the `t.test()` function will automatically calculate these numbers). Finally, your function should return an error when the user provides any plot option other than "boxplot" or "histogram". See the image KQ2_plot.png for examples of what the function should output.

```{r}
bigspotify <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv')

bigspotify

```

```{r}
#| error: true

test_genre_popularity <- function(df, genre1, genre2, plot) {
  
  smaller_genre <- df |>
    filter(playlist_genre %in% c(genre1, genre2))
  
  test <- t.test(track_popularity ~ playlist_genre, data = smaller_genre)
  
  if (plot == "boxplot") {
    
    ggplot(smaller_genre, aes(x = playlist_genre, y = track_popularity)) +
      geom_boxplot() +
      labs(
        title = paste('Comparing "', genre1, '" and "', genre2, '" genres by popularity', sep = ""),
        subtitle = paste(
          "Two sample t-test: T =",
          round(test$statistic, 2),
          "and p-value =",
          round(test$p.value, 4)
        )
      )
    
  } else if (plot == "histogram") {
    
    ggplot(smaller_genre, aes(x = track_popularity)) +
      geom_histogram() +
      labs(
        title = paste('Comparing "', genre1, '" and "', genre2, '" genres by popularity', sep = ""),
        subtitle = paste(
          "Two sample t-test: T =",
          round(test$statistic, 2),
          "and p-value =",
          round(test$p.value, 4)
        )
      )
    
  } else {
    stop(" Use boxplot or histogram as the plotting option")
  }
}
# This should work
test_genre_popularity(df = bigspotify, 
                      genre1 = "rap", 
                      genre2 = "rock", 
                      plot = "boxplot")

# This should prompt an error message that says: 
#   "Please use boxplot or histogram as the plotting option"
test_genre_popularity(df = bigspotify, 
                      genre1 = "rap", 
                      genre2 = "rock", 
                      plot = "density")


# NOTE: the only genre options are pop, rap, rock, latin, r&b, and edm
#  don't worry about an error message if one of these is not used
```

3. One of the harder concepts in statistics is the understanding of what the percentage in a confidence interval really means. Many introductory students believe that 95% confidence is a statement about the likelihood of the population parameter to be within the interval bounds. While not totally off-base, a better way to think about this is that if we take 1000 of different samples we expect around 950 of those samples to create 95% CIs that contain the true population parameter. Let's explore this.

Write a for loop that takes 10,000 samples of size 45 from a normal distribution with a true population mean of 25 and a true standard deviation of 4. Then calculate the lower bound and upper bound of a 95% confidence interval for the mean using each sample. Finally, check to see what proportion of simulated samples actually contain 25. (Spoiler - it should be close to 0.95)

Here are a couple hints and reminders:

- `rnorm(n = 20, mean = 0, sd = 1)` will generate a random sample of size 20 from a Normal distribution with mean 0 and standard deviation 1.
- The equation for a 95% confidence interval is the sample mean plus/minus 1.96 times sample standard deviation divided by the square root of sample size.

$$\bar{x} \pm 1.96 \times \frac{s}{\sqrt{n}}$$
- The mean of a vector of logicals will give you a proportion.



```{r}
# step 1 : I will set parameters
sample_size <- 45
num_samples <- 10000
true_mean <- 25
true_sd <- 4

# step 2: I will store whether each CI contains the true mean
yes_contains_mean <- vector("logical", num_samples)

# create loop
for (i in 1:num_samples) {
  samp <- rnorm(n = sample_size, mean = true_mean, sd = true_sd)  # To generate a sample
  xbar <- mean(samp) # to compute sample statistics
  s <- sd(samp)
  

  lower <- xbar - 1.96 * (s / sqrt(sample_size))   # to compute CI bounds
  upper <- xbar + 1.96 * (s / sqrt(sample_size))
  
  # check if CI contains true mean
  yes_contains_mean[i] <- (lower <= true_mean & upper >= true_mean)   # to check if CI contains true mean
}


mean(yes_contains_mean) # to calculate proportion

```

Bonus: Turn your for loop from part a) into a function with inputs for `sample_size`, `num_samples`, `true_mean`, and `true_sd`. 

```{r}
ci <- function(sample_size, num_samples, true_mean, true_sd){
  yes_contains_mean <- vector("logical", num_samples)

# create loop
for (i in 1:num_samples) {
  samp <- rnorm(n = sample_size, mean = true_mean, sd = true_sd)  # To generate a sample
  xbar <- mean(samp) # to compute sample statistics
  s <- sd(samp)
  

  lower <- xbar - 1.96 * (s / sqrt(sample_size))   # to compute CI bounds
  upper <- xbar + 1.96 * (s / sqrt(sample_size))
  
  # check if CI contains true mean
  yes_contains_mean[i] <- (lower <= true_mean & upper >= true_mean)   # to check if CI contains true mean
}


mean(yes_contains_mean) # to calculate proportion

}

ci(45, 10000, 25, 4)
# change output to test 
ci(45, 10000, 25,15)
```

